### 2011-12-13 report/paper.tex
Added final changes and fixed hyperref version dependency

% Matthew Monaco
% Andy Sayler
% Landon Spear
% University of Colorado
% Networked Character devices
% Fall 2011

\documentclass[11pt,twocolumn]{article}

\usepackage[text={6.5in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}

\bibliographystyle{plain}

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\tiny,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=none,
  frameround=tttt
}

\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\newenvironment{packed_enum}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\begin{document}

\title{Networked Character Devices}
\author{Monaco, Matthew \and Sayler, Andrew \and Spear, Landon \\
  \and University of Colorado \\
  \texttt{first.last@colorado.edu}
}
\date{\today}

\maketitle

\begin{abstract}
While there are various projects which provide a network abstraction layer for
specific high-level subsystems in the Linux kernel, there aren't any low-level
general purpose solutions. Network character devices (NCD) is a system for
exporting character device interfaces over a network. This capability
divorces the traditional requirement that a device can only be used by
the host to which it is physically connected in favor of a model that
allows a device to be connected to one host but used by another.
In addition to networking the character device interface for which it
is optimized, NCD also works with regular files, block devices,
sockets, and pipes. We demonstrate how our NCD system works, as well
as a few examples of the unique systems that can be built atop the NCD
interface.
\end{abstract}

% Landon Sections

\section{Introduction}
\label{sec:introduction}

We have designed and implemented Network Character Devices (NCD) to
enable a new approach to using and sharing devices. Today, networking 
technology is becoming faster and more reliable at incredible rates. 
Functionality that has been traditionally believed to only be 
feasible at the local level can now be pushed to remote
machines. Our project aims to allow 
a character device present on a server to be exported to
a client over a network connection. Many systems use character device
interfaces, from Web cams, mice and monitors to USB flash drives and 
external hard drives. NCD works by abstracting the character device interface 
over IP. In Unix-like operating system, the character device interface
is a file interface. While designed and optimized for
character devices, this generality actually allows the NCD system to export 
any file. Being able to dynamically export devices over network
connections suggests a new cloud-like paradigm of device usage where 
locality is required only for reduced latency, which may not be a
concern for certain devices.

When a device is exported via the NCD system we would like it to give the
appearance of having essentially ``moved'' the device to the client 
by forbidding user level accesses on the server.
Acquiring a remote device will be
done through an advertisement protocol that exists in user space so
that details may be supplied about the server and how the connection
should be made (e.g. UDP vs TCP).  The client-side NCD's functionality
exists in kernel space so as to maintain transparency to
user level operations. Once imported, a device should be accessed just
like a normal local device.  This is achieved by the NCD kernel module
intercepting i/o calls and sending them out over the network to the
server. On the server, a userspace daemon accesses the physical
device to emulate the client's activity. The overall structure of the
NCD system is present in Figure \ref{fig:system-full}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{system-full-g.png}
  \caption{The NCD system. See Figure \ref{fig:system-working}
    for currently implemented system state.}
  \label{fig:system-full}
\end{figure}

Currently we have achieved a successful implementation of the low
level functionality for NCD. The client uses a kernel module written
for Linux 2.6.x+ while the server uses a userspace daemon. NCD is
still rough and in need of further development, as it tends to be
crash prone if not treated with care. The groundwork has been laid for
exporting any device, however many device subsystems access their
devices by avoiding the character device interface and these must be
dealt with on an individual basis. Section \ref{sec:futurework}
details some of these challenges and our plans for their resolution.

\section{Related Work}
\label{sec:relatedwork}

There are many other projects in existence that implement similar
network sharing in one form or another that provided inspiration for
NCD.

NFS was the first implementation of generalized file sharing
over remote network connections.  However, NFS only supports normal
files and cannot interface directly with devices \cite{nfs-source}.

NBD (Network Block Device) was rolled into the Linux kernel starting
with version 2.1.101. It enables users to export block devices over the
network, allowing low level access to disks and other block devices.
However, NBD provides no
exclusive access guarantees and users must ascertain on their own
whether the remote drive is being used. It also mandates buffered
operation and is optimized for block, not character, device operation
\cite{nbd-source}.

CUSE (Character Device in User Space) is an extension of FUSE that
allows bringing a character device into a FUSE filesystem.  This allows
users to access the device directly and potentially export it using
user-level networking \cite{cuse-article}. We considered using CUSE in
our NCD implementation, but opted for a direct in-kernel implementation
instead since it is more standard.

USB/IP is a project to share USB devices over IP networks. This project is
very similar to NCD (as USB devices are character devices) but is
tailored specifically to allow sharing USB devices over a
network connection by intercepting USB specific data and encapsulating
it in IP \cite{usbip-source}. It does not support a more generalized
notion of general character device networking.

Plan 9 is a distributed operating system developed as a successor to
Unix by Bell Labs.  It implicitly allows sharing devices over a
network via file interfaces. However, Plan 9 does not currently see
wide deployment \cite{plan9-pike}. We hope to bring the NCD concept to
the more widely deployed Linux environment.

% Matt Sections

\section{Architecture}
\label{sec:architecture}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{arch-03-g.png}
  \caption{The basic client-server model of NCD.}
  \label{fig:architecture}
\end{figure}

On a very high level, NCD uses a basic client server model. The client
issues a request to the server, which in turn issues a reply. The
server never initiates any communication on its own.

\subsection{Client}

The client is currently implemented entirely within an out-of-tree kernel
module. This is arguably not the most robust, nor secure, solution. See
Section \ref{sec:futurework} for details on splitting out some functionality
into a user-space helper.

Devices which have been imported on the client reside within the
devtmpfs file system. This is almost always mounted on \texttt{/dev}. By
default an imported device will take on the name
\texttt{/dev/netchar/importN}, where \texttt{N} is the next available
number. In the future (and without much effort), imported devices will
be able to take on their original name from the server. For example, an
input device should be located at \texttt{/dev/input/eventN}. Furthermore,
the user or administrator may choose any mapping (s)he see fits.

\subsection{Server}

The server is a typical user-space daemon. It
currently only serves a single file. Whether or not the server should
handle multiple exports is still up for evaluation.

The server listens on a port specified on the command line. There is no
default. When a request is received from the client, it will correspond
to one of the standard VFS file operations. Currently, \texttt{open()},
\texttt{close()}
\footnote{The counterpart for \texttt{close()} is actually referred to
as \texttt{release()} within the kernel.}
, \texttt{read()}, and \texttt{write()} are supported.

After receiving a request, the server then performs the requested
operation locally and returns the result to the client in a reply
message. For example, if the client sends an \texttt{open()} request,
the server will open and hold a file descriptor on the client's behalf.

This straightforward architecture will actually support any file type
(with certain restrictions for each type). Character devices, regular
files, sockets, and pipes should work fully (when all file operations
are implemented). Block devices and even directories will work to some
extent.

\subsection{Protocol}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{proto-g.png}
  \caption{NCD operates over a simple request-reply protocol. Bulk data
    immediately follows formal messages, as required.}
  \label{fig:protocol}
\end{figure}

The protocol is implemented as a simple request-reply system over
\texttt{INET} sockets. All request types are encapsulated within a
single request structure (see Section \ref{sec:implementation}).
Similarly, there is only a single reply structure to account for all
reply types.

Bulk data is sent in the raw immediately following the corresponding
request/reply. In the case of a \texttt{read()}, the targeted data
immediately follows the reply. In the case of a \texttt{write()}, the
targeted data immediately follows the initial request.

\section{Implementation}
\label{sec:implementation}

\subsection{Protocol}

\lstinputlisting[
  caption={The structures which make up the NCD protocol.},
  label=lst:protocol
]{protocol.c}

The protocol for NCD operates over TCP sockets. Despite the potential
for large amounts of streaming data over the Internet --- data which may
be acceptable to drop --- some messages are unacceptable to lose. Any
request or reply should be guaranteed. The data following a
\texttt{read()} or \texttt{write()} (figure \ref{fig:protocol}), for
example, may not be as important. Therefore, in the future it may be
beneficial to implement the core protocol over TCP and bulk data over
UDP (section \ref{sec:futurework}).

Because there is only one type of packet for a request, and another for
a reply, (Listing \ref{lst:protocol}) there is a small amount of wasted
space on some file operations. We have done our best to reduce this
overhead as much as possible, making use of a cascade of nested unions
and structs.

The request struct is suitable to handle all of the function arguments
for all of the operations we support. The reply struct is much simpler
as functions only have a single return value. While some file operations
have output parameters and some userspace system calls do as well, there
aren't any cases where these need to be passed over the network.

\subsection{Client}

\lstinputlisting[
  caption={The NCD kernel module has a number of roles.},
  label=lst:client
]{cdev-driver.c}

The client resides entirely within a kernel module \footnote{The merits
of including everything within a kernel module are discussed in section
\ref{sec:futurework}.}. A typical userspace program can be sloppy
with cleanup because all memory is freed on termination. In contrast, a
kernel module must be very careful because its lifetime is shorter than
the kernel itself. Therefore it is very import to correctly handle
errors and cleanup all memory in the \texttt{module\_init} and
\texttt{module\_exit} routines.

Some of the basic features of the NCD kernel module are listed in
Listing \ref{lst:client}. NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

The \texttt{struct class} is somewhat synonymous with a kernel
subsystem. Because the major/minor number is dynamic, it is with this
class that userspace is able to persistently refer to NCD character
device nodes (see \ref{subsec:udev}).

\texttt{Struct file\_operations} is used to define the handlers for all
file operations which we support for our character devices \footnote{See
\texttt{linux/fs.h} for a complete list of file operations.}.

Most importantly, \texttt{struct cdev} ties together the major/minor
number with the given set of file operations. With this it is possible
for a kernel module to register for many major/minor numbers (typically
a contiguous range), while having fine-grained control over which
subsets are handled by which distinct groups of file operations. This
will be important in the future as specific subsystems are handled
differently by NCD to increase functionality.

Finally, \texttt{struct device} is used to actually create a character
device node under \texttt{/dev} in userspace. Unfortunately, not
everything we might want to do to this device can be accomplished in the
kernel (permissions for example).

\subsection{Server}

\lstinputlisting[
  caption={A code snippet from the server illustrating a \texttt{read()}
  file operation},
  label=lst:server
]{server.c}

The server will spend most of its time sitting on a blocked read of the
socket file descriptor. Once the client sends a request, it will be
available to the server in the appropriate structure.

In the snippet given in Listing \ref{lst:server}, a \texttt{read()} file
operation is handled. A buffer must be allocated with a size
indicated by the original request. Next, the server can perform the
corresponding operation and send a reply message to the client. Finally,
the bulk data is sent to the client. All buffering is performed by the
underlying \texttt{SOCK\_STREAM} and TCP protocols.

The server itself is, and will always be, simple compared to the
in-kernel module. As NCD grows, the server's functionality should remain
fairly static. Section \ref{sec:futurework} discusses some enhancements
which will make the server more stable, however it doesn't require many
new features.

\subsection{Misc}

\subsubsection{udev}
\label{subsec:udev}

\lstinputlisting[
  lastline=1,
  caption={A simple \texttt{udev} rule for managing NCD device
  permissions.},
  label=lst:udevrule
]{udev.txt}

A lot of magic is performed on character devices in userspace by
the Linux \texttt{udev} system.
Currently we are using \texttt{udev} to manage the
permissions of imported devices (Listing \ref{lst:udevrule}).

\lstinputlisting[
  firstline=3,
  caption={\texttt{udev} attributes required by userspace programs such
  as the \texttt{Xserver} for knowing a device is a keyboard.},
  label=lst:udevattr
]{udev.txt}

For more functionality, imported character devices must have certain
attributes such as those for the keyboard in Listing \ref{lst:udevattr}.
Unfortunately setting them retroactively does not convince all programs
of a particular devices type. Therefore interfacing with specific kernel
subsystems such as \texttt{input} and \texttt{usb} will be required in
the future for enhanced functionality.

\subsubsection{ncadmin}

\lstinputlisting[
  language={},
  caption={The potential management interface for the NCD system.},
  label=lst:ncadmin
]{ncadmin.txt}

As NCD grows, it will be important for the client to be able to handle
dynamically added and removed devices. Listing \ref{lst:ncadmin}
provides an idea of what an interface would look like to provide such
functionality. This would give NCD the ability to create its character
devices anywhere under \texttt{/dev}; they could be placed in a similar
location to the corresponding device on the server.

% Andy Sections

\section{Results and Evaluation}
\label{sec:results}

The Network Character Devices system is still a work in
progress. We will need to expand the system beyond it's current state
before the NCD concept is ready for production-level use.

None the less, we have gained many insights into the advantages such a
system might provide. We have also encountered the difficulties
implementing such a system into an existing OS like Linux entails.

\subsection{Current Status}
\label{sec:currentstatus}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{system-working-g.png}
  \caption{The NCD system with currently functioning portions
    highlighted.}
  \label{fig:system-working}
\end{figure}

As of this writing, we have completed a partial implementation of
the Network Character Devices system (Figure
\ref{fig:system-working}). Current functionality includes:

\begin{packed_item}
\item Support for exporting a single character device from the server
\item Support for importing a single character device on the client
\item Support for \texttt{open()}, \texttt{close()}, \texttt{read()},
  and \texttt{write()} calls
\item Basic Linux \texttt{udev} support for automatic node
  provisioning on the client
\end{packed_item}

This set of core functionality provides a working proof-of-concept
Network Character Device system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

It should be noted, however, that some work is still required before
the Network Character device system can be considered a production
ready utility. Most notably, the following features are necessary for
a fully functional NCD system, but are not yet included in our
implementation:

\begin{packed_item}
\item Multi-device, multi-server client-side import support
\item Multi-device server-side export support
\item Support for exclusive use and protection of exported devices
\item Support for ioctl calls
\item Support for providing and obtaining metadata for exported
  devices
\item Support for ``advertisement'' of available exported devices
\item Support for integrating imported devices into other kernel
  subsystems (human interface subsystem, audio subsystem, etc)
\item Addition of UDP and TCP support with user selection
\item Addition of graceful connect and disconnect handling in real-time
\end{packed_item}

Section \ref{sec:futurework} provides some insight into the addition
of some of these expanded features.

\subsection{Advantages}
\label{sec:advantages}

The primary advantages of using Networked Charterer Devices are:

\begin{packed_item}
\item Divorces peripherals from the computers to which they are
  attached
\item Allows simple creation of a ``Device Server''
\item Extends the ``Everything Is A File'' -nix philosophy
\item Allows more efficient use of devices
\item Allows devices to be networked in a generalized, uniform manner
\end{packed_item}

By allowing the interface for any device to be exported over the
network, the NCD system effectively removes the traditional restriction
that devices are only available on the hosts to which they are
physically connected. Removing this restriction has far reaching
implications (see Section \ref{sec:potentialuses}). One of which is
increased efficiency of device utilization. By allowing devices to
be exported from machines where they are not needed and imported on
machines where they are, we can insure that devices are efficiently
utilized without having to worry about the underlying physical locations
of the devices.

One of the foremost tenants of the Unix philosophy is the ``Everything
is a file'' concept. The NCD system extends this concept to
``Everything is a networked file''. This allows us to create ``Device
Servers'' as easily as we currently create file servers. The NCD
system allows us to network any device via the general and
well-understood file interface instead of relying on devices specific
networking implementations.

\subsection{Challenges}
\label{sec:challenges}

While Network Charter Devices provide a number of advantages, we have
also stumbled upon a number of challenges when it comes to properly
building and utilizing an NCD system. In particular, the NCD system:

\begin{packed_item}
\item Does not work well when Character Device file is not the exclusive
  device interface
\item Requires exported devices to be ``de-integrated'' from host system
\item Requires QOS guarantees and high quality network connection
\end{packed_item}

The original intention of the Unix and Linux \texttt{dev} file
system and the corresponding Charter Device file interface was to
provide the sole interface to device drivers. Today, while Character
Device file interfaces are still is use, they are no longer the sole
devise driver interface. Auxiliary systems like procfs
\cite{interfaces}, sysfs \cite{sysfs-mochel}, and netlink sockets
\cite{netlink-ayuso} provide supplementary and/or replacement interfaces
for the traditional Character Device file interface. Furthermore, some
device drivers interact directly with in-kernel subsystems (HID subsystem,
etc), providing no direct userspace interface at all. Since our system
focuses on abstracting the Charter Device file interface over the
network, use of these additional or alternate interfaces make devices
incompatible with the NCD system. This incompatibility can be mitigated
by extending the NCD system to include these alternate interfaces, but the
breakdown of the original Unix ideal of having a single device driver
interface makes building systems like NCD much more difficult and less
generalized.

Most modern Operating Systems, including Linux, assume that any device
physically connected to the system is available for the system to
use. As mentioned previously, Network Character Devices break down this
assumption. Thus, one of the challenges involved in implementing an NCD
system is to remove the OS's assumption that any connected device is
available for local use. Modern-day plug-and-play auto-configuration
systems make this task even more difficult, and the means for breaking
down the availability assumption tend to be unique to a given
OS. Effectively removing this assumption is one of the larger,
still-open, NCD challenges.

Finally, since the NCD system makes a networked device appear as
though it is connected locally, the NCD system does not gracefully
handle situations that would never occur locally like network faults
and errors. Thus, the NCD system performs best when running on a
network with QOS guarantees and the necessary throughput to support all
NCD traffic.

\section{Future Work}
\label{sec:futurework}

As Figure \ref{fig:system-working} shows, there are still several
portions of the NCD system that need implemented before the system is
ready for production-level use. In particular, the top priorities
currently are:

\begin{packed_item}
\item Add multi-device, multi-server support
\item Add kernel sub-system handler layer (HID, etc)
\item Add device protection and removal capabilities to server system
\item Add advertisement protocol and unreliable data protocol option
\item Separate networking and memory-intensive code from client kernel module
\end{packed_item}

The current NCD system only supports a single device per server and per
client. To be truly useful and to enable the construction of the
kinds of system we envision NCD being used for (Section
\ref{sec:potentialuses}), we must at least
expand the client-side
system to support importing an arbitrary number of devices from one
or more servers. Ideally, servers should also be capable of serving
multiple devices, but this is of less importance than multi-device
client handling.

As we mentioned in Section \ref{sec:challenges}, some modern device
drivers bypass the traditional Character Device file interface. Thus, we
must expand the NCD system to accommodate the additional interfaces and
subsystems that are commonly in use. This includes special import
handling for in-kernel subsystems like the X Server and HID
subsystems, as well as generalized support for sysfs and procfs import
and export.

We also plan to add a means for device protection and exclusion to
the NCD system. This will include both the need to remove exported
devices from use by local server-side operating systems as well as controlling
the number of clients that can connect to a single exported
device. Controlling the latter point will require a per-device
definition, probably provided as part of the export configuration, of
how many clients may simultaneously utilize each device. Some devices will
require exclusive access while other may be capable of handling read
and/or write request from multiple clients simultaneously.

Currently, all NCD traffic is sent via TCP. While control messages
should always use TCP to insure system state and stability, there are
device types who generate data more appropriate for UDP operation (web
cams, etc). In the future, we hope to allow the user to specify
whether the data traffic for a given device will be delivered via TCP
or whether it can be delivered using UDP.

Since large scale NCD deployments may include thousands of exported
devices, the NCD system requires the addition of an advertisement and
discovery protocol so that client may ascertain what resources are
currently available for import. Such a system will need to provide
information about what devices are available, as well as metadata
describing the basic characteristics of a given device.

Finally, the current client-side NCD implementation performs both direct
network access and memory-intensive processing directly in the kernel
code. This approach was taken in our initial implementation to control
complexity.
For security reasons, it is desirable to move the direct
network communication from the kernel code to user-space code where
the repercussions of a security breach as less extreme.
For performance reasons, it is desirable to move memory
intensive operations from the kernel code (where they are limited to
the available kernel memory space) to user-space code where the memory
availability is far greater and the use restrictions more relaxed. To
solve these issues, we
foresee building a client-daemon that will handle the bulk to the
client-side operations and that will, in turn, communicate directly with
a more minimal client kernel module only when necessary.

\section{Potential Uses}
\label{sec:potentialuses}

The shift from the traditional concept of locally available devices to a
broader concept of generalized network accessible devices has the
potential to allow the creation of a variety of novel
applications. Here we explore a few of the possible applications of an
NCD system.

\subsection{Networked KVM System}
\label{sec:networkedkvm}

KVM (Keyboard-Video-Mouse) systems are commonly used in data centers
and other areas where a user wishes to use a single keyboard, video,
and mouse to control multiple machines. KVM systems range from simple KVM
switches that allow the user to select which local machine is
physically connected to his or her i/o devices to
networked KVM systems that allow the user to use a
local KVM setup to interface with a computer located elsewhere on the
network. The latter class of networked KVM devices traditionally require
expensive propriety hardware.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{kvm-g.png}
  \caption{A networked KVM system built using NCD.}
  \label{fig:kvm}
\end{figure}

As Figure \ref{fig:kvm} shows, we can use Networked Character Devices
to quickly and easily build a networked KVM system in software,
forgoing the need for expensive specialized hardware. The user simply
connects a spare mouse and keyboard to an NCD server, tells the NCD
system to export these devices, and then imports them on the client
machine (s)he wishes to control. To the client, the devices appear as
a locally connected keyboard and mouse, and the user may issue input to
the client machine via the keyboard and mouse connected to the
server.

As we mentioned previously, an NCD based KVM system would require the
appropriate import interface to the Linux HID subsystem, but once this is
in place, building a networked KVM atop the NCD system would be a
rather trivial matter.

\subsection{Distributed RNG System}
\label{sec:distributedrng}

Another example of the novel systems that can be built using Network
Charter Devices is a distributed random number generator. Linux has a
fairly robust random number generator built into it and accessible as
a Character Device at \texttt{/dev/random}. This RNG relies on the
local entropy pool (generally filled by human interaction, interrupt
timings, and other high entropy events) to generate random
numbers. Thus, the speed at which random numbers can be generated is
limited on any single system to the locally available
entropy. This can severely hinder the
performance of systems that require random numbers\footnote{Linux also
provides the \texttt{/dev/urandom} RNG which is a pseudorandom number
generator seeded off the true
\texttt{/dev/random}. \texttt{/dev/urandom} trades ``randomness'' in
favor of speed. It is thus handy when one just needs to quickly
generate pseudorandom numbers, but is not appropriate for cryptography or
other secure random system.}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{rng-g.png}
  \caption{A distributed Random Number Generator system built using NCD.}
  \label{fig:rng}
\end{figure}

We can imagine a system that uses Network Character Devices to overcome
the normal limitations related to relying on locally available entropy. This
system would be capable of generating cryptographically secure
randomness while also delivering high performance.

Figure \ref{fig:rng} shows how such a system might work. In this
distributed RNG system, multiple hosts act as NCD servers by exporting
their local \texttt{/dev/random} character devices. These devices are
all imported by a single client running some from on RNG aggregator
that combines the relatively slow output from the local RNGs into a
single high speed RNG output. We can effectively pool the entropy
available on numerous hosts to drive a single high performance RNG
system.

As in the KVM system, using Networked Character Devices makes the
implementation of such a system relatively straight forward.

\section{Conclusion}
\label{sec:conclusion}

Network Character Devices provide an extension of the standard device
driver file interface across the network. They allow for device
capabilities to be exported from where they are not needed and imported
where they are. By separating device usage from the device's physical
location and host, the NCD system provides a generalized, transparent,
means for abstracting device functionality across a network.

As the potential uses demonstrate, the power of
the NCD concept is in the ability to achieve complex solutions without
needing to resort to specialized device networking protocols and other
complex ad-hoc systems. We hope to continue the development of the NCD
concept and to utilize it to build powerful, but easily implementable,
device-dependent systems.

\nocite{*}
\bibliography{refs}

\end{document}

% vim: set ts=2 sts=2 sw=2 et spell tw=72 : %